---
title: "Figure 3"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez-Otero, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

Panel A of figure 2 is isothermal titration calorimetry data that was analyzed outside of R. 

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(modelr)
library(broom)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../tools/plotting_tools.R")

# Modify the plot theme
theme_set(theme_notebook())
```

# Results

Here's the info for each well:
```{r}
metadata <- read_csv("../../../data/Spectroscopy/2019_10_18_ET_ctDNA_metadata.csv")

metadata %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')

```

## Abs Spectrum 1

Now let's read in the first absorbance spectra taken for each condition and we'll add the appropriate metadata:
```{r}

df <- read_csv("../../../data/Spectroscopy/2019_10_18_ET_ctDNA_abs_spectra_1.csv") %>%
    gather(key = 'well', value = 'intensity', -Wavelength)

df_meta <- left_join(df, metadata, by = 'well')

df_meta %>% kable %>% kable_styling() %>% scroll_box(height = '400px')

```

Let's look at all the data:

```{r}
#ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + geom_smooth(se = T)

ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + geom_point()

```

This looks pretty much like the past notebook with interphenazine ET. If we look out at 690 for the PYO abs peak it looks like the orange PYOox is higher than all the other conditions. Let's zoom into that region for a closer look:

```{r}
ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + 
  geom_jitter(shape=21)+coord_cartesian(xlim = c(600,700), ylim = c(0,0.2))
```

Ok, this looks pretty much as we expected. All of the conditions show 690 abs closer to PYOred than PYOox.

Let's try and condense this info into some pointranges:

```{r}
df_690 <- df_meta %>% group_by(red_ox, Wavelength, DNA) %>% mutate(mean = mean(intensity), sd = sd(intensity))  %>% filter(Wavelength == 690)

ggplot(df_690, aes(x = red_ox, y = intensity, color = DNA )) + 
  geom_jitter(shape = 21, width = 0.1) + 
  geom_pointrange(data = . %>% filter(rep ==1), 
                  aes(y = mean,ymin = mean-sd,ymax = mean+sd), 
                  position = position_jitter(width = 0.1))
```


```{r}


labels = labels = c('PBS + PYO[ox]','PYO[red] + PBS','PCA[red] + PYO[ox]','PCN[red] + PYO[ox]')


plot_ctDNA_690 <- ggplot(df_690, aes(x = red_ox, y = mean, color = DNA)) +  
  geom_hline(yintercept = 0.033, linetype = 2, color = 'light gray')+
  geom_jitter(aes(y = intensity), shape = 21, position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.1, jitter.height = 0))+
  geom_pointrange(aes(ymin = mean - sd, ymax = mean + sd), position = position_dodge(width = 0.4), size = 0.5, fatten = 0.75) + 
  scale_x_discrete(limits = c('PBS_PYO','PYO_PBS','PCA_PYO','PCN_PYO'), labels = parse(text = labels )) + 
  scale_color_manual(breaks = c(F,T), labels = c('-DNA','+DNA'), values = colorblind_palette)+
  labs(x = NULL, y = 'Absorbance at 690nm', color = NULL) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.5,1))

plot_ctDNA_690
```

Ok these pointranges show mean +/- 2 standard deviations. So it looks like for all of the conditions there's no significant difference +/- ctDNA.

## Abs690 Kinetic

The Spectrum data looks good, let's also take a look at the kinetic data at 690 to see if there's a difference in the reaction rates. First we'll read it in:

```{r}

df_kin <- read_csv("../../../data/Spectroscopy/2019_10_18_ET_ctDNA_abs_690kinetic_1.csv") %>%
    gather(key = 'well', value = 'intensity', -min)

df_kin_meta <- left_join(df_kin, metadata, by = 'well')

df_kin_meta %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')

```

Now let's look at the values over time (~2min)

```{r}

ggplot(df_kin_meta, aes(x = min, y = intensity, color = DNA)) + geom_point() + facet_wrap(~red_ox)

```

Ok, the intensity trends we observed with the spectra looks the same here. There are no obvious differences in slope, so I still don't think we can say anything about reaction rates...they all proceed too fast.

```{r}

theme_set(theme_figure())

fig_3 <- plot_grid(NULL, plot_ctDNA_690, ncol = 2, scale = 0.95, labels = 'AUTO', label_size = 12, rel_widths = c(1,1.5))

fig_3

save_plot("../../../figures/phz2019_Fig_3.pdf", fig_3, base_height = 2, base_width = 3.4)
```

------

# Adding DNA CT to same figure

-----

# Panel E - MM Discrimination

This data is electrochemical and contains the third/final sweep of an electrode with two different DNA monolayers.
```{r}
df_noO2 <- read_csv('../../../data/Electrochemistry/DNA_modified_electrode_noO2.csv',comment = "#") %>% 
  mutate(Construct=fct_relevel(Construct,'wm','mm'))

df_noO2 %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')
```

Let's make the figure:
```{r}

plot_dnaCT_noO2 <- ggplot(df_noO2,aes(x=E,y=Current, color = Construct))+
  geom_path()

plot_dnaCT_noO2_styled <- plot_dnaCT_noO2 +
  scale_x_reverse(labels = mV_label)+
  scale_y_continuous(labels = nA_label)+
  scale_color_manual(breaks = c('wm','mm'), labels=c("Well matched DNA","Mismatched DNA"), values = c("#66CCFF","#FFCC66")) +
  labs(x='E (mV vs. Ag/AgCl)',y='Current (nA)')+
  theme(legend.position = c(0.3, 0.8),
        legend.title = element_blank(),
        legend.background = element_rect(fill=NA))

plot_dnaCT_noO2_styled
```




# Panel F - DNA CT Catalytic

```{r}
df_wO2 <- read_csv('../../../data/Electrochemistry/DNA_modified_electrode_wO2.csv',comment = "#") %>% 
  mutate(Condition=paste(Construct,Aerobic,PCN,sep='_')) %>% 
  mutate(Condition=fct_relevel(Condition,"wm_TRUE_TRUE","mm_TRUE_TRUE","wm_TRUE_FALSE"))

df_wO2 %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')
```

```{r}
plot_dnaCT_wO2 <- ggplot(df_wO2 %>% filter(Aerobic==T),aes(x = E, y = Current, color = Condition))+
  geom_path()

plot_dnaCT_wO2_styled <- plot_dnaCT_wO2 +
  scale_x_reverse(labels = mV_label)+
  scale_y_continuous(labels = nA_label)+
  scale_color_manual(
    breaks = c('wm_TRUE_TRUE','mm_TRUE_TRUE','wm_TRUE_FALSE'), values = c("#66CCFF","#FFCC66","gray"),
    labels=c(expression("wmDNA"+"PCN"+O[2]),expression("mmDNA"+"PCN"+O[2]),expression("wmDNA"-"PCN"+O[2]))
    ) +
  labs(x='E (mV vs. Ag/AgCl)',y='Current (nA)')+
  theme(
    legend.position = c(0.3, 0.8),
    legend.title = element_blank(),
    legend.background = element_rect(fill=NA))

plot_dnaCT_wO2_styled
```

# Spectroscopy

```{r}


df_spec <- read_csv("../../../data/Spectroscopy/2019_08_29_spectroscopy_RuRh_dphz.csv") %>% filter(exp_num == 2)

df_spec <- df_spec %>% 
  mutate(min = ifelse(ruphen_conc==0,intensity, NA)) %>% 
  group_by(time) %>% 
  mutate(bg = min(min, na.rm = T)) %>% 
  mutate(bg_sub = intensity - bg)

plot_raw_spec <- ggplot(df_spec %>% filter(run <= 6), aes(x = time, y = intensity, color = quencher_eq, group = quencher_eq)) + 
  geom_path() + 
  scale_x_continuous(limits = c(NA, 5e-7), labels = ns_label) +
  labs(x = 'Time', y = 'Intensity (620nm)', color = 'Quencher\nEquivalents') + 
  guides(color = F)
  
plot_raw_spec

df_liq <- read_csv("../../../data/Spectroscopy/2019_08_29_spectroscopy_RuRh_dphz.csv") %>% filter(exp_num == 7)

plot_liq_spec <- ggplot(df_liq %>% filter(run <= 6), aes(x = time, y = intensity, color = quencher_eq, group = quencher_eq)) + 
  geom_path() + 
  scale_x_continuous(limits = c(NA, 5e-7), labels = ns_label) + 
  labs(x = 'Time', y = 'Intensity (620nm)', color = 'Quencher\nEquivalents') + guides(color = F)

plot_liq_spec

ggplot(df_spec %>% filter(run %in% 2:6) %>% filter(time>30e-9), aes(x = time, y = bg_sub, color = quencher_eq, group = quencher_eq)) + geom_path()

df_fac <- bind_rows(df_spec, df_liq)

labels <- c('2' = "biofilm", '7' = "liquid")

df_fac$exp_num <- factor(df_fac$exp_num, levels = c('7','2'))

plot_fac <- ggplot(df_fac %>% filter(run <= 6), aes(x = time, y = intensity, color = quencher_eq, group = quencher_eq)) + 
  geom_path() + 
  scale_x_continuous(limits = c(NA, 5e-7), labels = nA_label) + 
  labs(x = 'Time (ns)', y = 'Intensity (620nm)', color = 'Quencher\nEquivalents') + guides(color = F) + facet_wrap(~exp_num, labeller=labeller(exp_num = labels), scale = 'free')

plot_fac

```

```{r}
df_spec_fit <- df_spec %>% 
  filter(run %in% 2:6) %>% 
  filter(time>30e-9) %>% 
  filter(bg_sub>0) %>% 
  group_by(run, quencher_eq) %>% 
  nest()

fit_biexp_bg_sub <- function(df){
  
  A1 = 0.02; lrc1 = 16.5; A2 = 0.02; lrc2 = 15.7
  
  mod <- nls(bg_sub~SSbiexp(time, A1, lrc1, A2, lrc2), data = df)
  
  mod
}

df_spec_models <- df_spec_fit %>% 
  mutate(models = map(data, fit_biexp_bg_sub)) 

df_spec_preds <- df_spec_models %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)

df_spec_ests <- df_spec_models %>% 
  mutate(ests = map(models, tidy)) %>% 
  unnest(ests)


df_spec_preds %>% head()
```


```{r}
#write_csv(df_exp_2_ests, "2019_09_27_spectroscopy_fits_bg_sub.csv")

df_spec_ests %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')
```


```{r}

ggplot(df_spec_preds, aes(x = time, group = run, color = quencher_eq)) + geom_path(aes(y = pred), color = 'black') + geom_point(aes(y = bg_sub),alpha = 0.2, shape = 21) +xlim(0,5e-7) + facet_wrap(~quencher_eq)

```


```{r}

plot_bg_spec <- ggplot(df_spec %>% filter(run %in% 1:6), aes(x = time, group = run, color = quencher_eq)) + 
  geom_point( aes(y = bg_sub),alpha = 0.2, shape = 21, size = 0.5) + 
  geom_path(data = df_spec_preds,aes(y = pred)) + scale_x_continuous(limits = c(NA,1e-6), labels = nA_label, name = 'Time (ns)') + 
  ylim(-0.001, NA) + labs(y = 'Background subtracted - Intensity (620nm)', color = 'Quencher\nequivalents')+
  guides(color = guide_colorbar(barwidth = 0.5, barheight = 2.5)) + theme(legend.position = c(0.75,0.75))

plot_bg_spec

```


```{r}
background_ests <- df_spec_ests %>% filter(is.na(quencher_eq))

#background_ests

ggplot(df_spec_ests, aes(x = quencher_eq, y = estimate)) + 
  geom_hline(data = background_ests, aes(yintercept = estimate), linetype = 2, color = 'light gray') + 
  geom_pointrange(aes(ymin = estimate - 2*std.error, ymax = estimate + 2*std.error)) +
  facet_wrap(~term, scales = 'free')

```

```{r}
df_spec_tau <- df_spec_ests %>% filter(term %in% c('lrc1','lrc2')) %>% mutate(tau = 1/exp(estimate)) %>% mutate(tau_low = 1/exp(estimate + 2*std.error), tau_high = 1/exp(estimate - 2*std.error))


background_spec_tau <- df_spec_tau %>% filter(is.na(quencher_eq))

#background_ests

plot_spec_tau <- ggplot(df_spec_tau, aes(x = quencher_eq, y = tau)) + 
  geom_hline(data = background_spec_tau, aes(yintercept = tau), linetype = 1, color = 'light gray') + 
  geom_pointrange(aes(ymin = tau_low, ymax = tau_high)) +
  facet_wrap(~term, scales = 'free', labeller = labeller(term = c(lrc1 = 'Component 1', lrc2 = 'Component 2'))) + 
  scale_y_continuous(labels = ns_label, limits = c(0,NA))

plot_spec_tau
```

```{r}
theme_set(theme_figure())

fig_4 <- plot_grid(plot_dnaCT_noO2_styled, plot_dnaCT_noO2_styled, plot_dnaCT_wO2_styled,
          plot_fac, plot_fac, plot_bg_spec, 
          ncol = 3, scale = 0.95, align = 'hv', axis = 'tblr', labels = 'AUTO', label_size = 12)

fig_4

save_plot("../../../figures/phz2019_Fig_4.pdf", fig_4, base_height = 4, base_width = 7)
```

```{r}
fig_4 <- plot_grid(plot_ctDNA_690, plot_ctDNA_690, plot_ctDNA_690,
                   plot_dnaCT_noO2_styled, plot_dnaCT_noO2_styled, plot_dnaCT_wO2_styled,
          plot_fac, plot_fac, plot_bg_spec, 
          ncol = 3, scale = 0.95, align = 'hv', axis = 'tlr', labels = 'AUTO', label_size = 12)

fig_4

save_plot("../../../figures/phz2019_Fig_3combo.pdf", fig_4, base_height = 6, base_width = 7)
```

```{r}
sessionInfo()
```