---
title: "Figure 3"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez Otero, Scott A. Trammell, Eric D.A. Stemp, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

Panels A, D, and G of figure 3 are diagrams made in illustrator, so plots are replicated as placeholders in this R generated figure. 

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Check packages
source("../../tools/package_setup.R")

# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(modelr)
library(broom)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../tools/plotting_tools.R")

# Modify the plot theme
theme_set(theme_notebook())

# Load text file import tools
source("../../tools/text_file_import.R")
```

# Fig. 3B - Interphenazine electron transfer (anoxic)

This experiment was conducted in a plate reader - here's the info for each well:
```{r}
metadata <- read_csv("../../../data/Spectroscopy/2019_10_18_ET_ctDNA_metadata.csv")

metadata %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')

```

## Absorbance Spectra

Now let's read in the first absorbance spectra taken for each condition and we'll add the appropriate metadata:
```{r}

df <- read_csv("../../../data/Spectroscopy/2019_10_18_ET_ctDNA_abs_spectra_1.csv") %>%
    gather(key = 'well', value = 'intensity', -Wavelength)

df_meta <- left_join(df, metadata, by = 'well')

df_meta %>% kable %>% kable_styling() %>% scroll_box(height = '400px')

```

Let's look at all the data:

```{r}
#ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + geom_smooth(se = T)

ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + geom_point()

```

These match characteristic phenazine spectra. If we look out at 690 for the PYO abs peak it looks like the orange PYOox is higher than all the other conditions. Let's zoom into that region for a closer look:

```{r}
ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + 
  geom_jitter(shape=21)+coord_cartesian(xlim = c(600,700), ylim = c(0,0.2))
```

All of the conditions show 690 abs closer to PYOred than PYOox, except for the control condition that only contains oxidized PYO.

Let's try and condense this info into some pointranges:

```{r}
df_690 <- df_meta %>% 
  group_by(red_ox, Wavelength, DNA) %>% 
  mutate(mean = mean(intensity), sd = sd(intensity))  %>% 
  filter(Wavelength == 690)

labels = c('PBS + PYO[ox]','PYO[red] + PBS','PCA[red] + PYO[ox]','PCN[red] + PYO[ox]')


plot_ctDNA_690 <- ggplot(df_690, aes(x = red_ox, y = mean, color = DNA)) +  
  geom_hline(yintercept = 0.033, linetype = 2, color = 'light gray')+
  geom_jitter(aes(y = intensity, shape = DNA),color = 'black', size = 0.5, stroke = 0.2, 
              position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.1, jitter.height = 0))+
  geom_pointrange(aes(ymin = mean - sd, ymax = mean + sd), 
                  position = position_dodge(width = 0.4),size = 0.5, fatten = 0.5) + 
  scale_x_discrete(limits = c('PBS_PYO','PYO_PBS','PCA_PYO','PCN_PYO'), labels = parse(text = labels )) + 
  scale_color_manual(breaks = c(F,T), labels = c('-DNA','+DNA'), values = colorblind_palette)+
  labs(x = NULL, y = 'Absorbance at 690nm', color = NULL) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.5,1)) + 
  scale_shape_manual(values = c(21,22))

plot_ctDNA_690
```

Ok these pointranges show mean +/- standard deviations. So it looks like for all of the conditions there's no significant difference +/- ctDNA, but the reactions with reduced PCA and PCN yield reduced PYO (and therefore a reaction occured).

## Kinetics

In addition to spectra taken immediately following mixing, we also acquired time courses at 690nm to see if there's a difference in the reaction rates. First we'll read in the data:

```{r}

df_kin <- read_csv("../../../data/Spectroscopy/2019_10_18_ET_ctDNA_abs_690kinetic_1.csv") %>%
    gather(key = 'well', value = 'intensity', -min)

df_kin_meta <- left_join(df_kin, metadata, by = 'well')

df_kin_meta %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')

```

Now let's look at the values over time (~2min)

```{r}

ggplot(df_kin_meta, aes(x = min, y = intensity, color = DNA)) + geom_point() + facet_wrap(~red_ox)

```

Ok, the intensity trends we observed with the spectra looks the same here. There are no obvious differences in slope, so I still don't think we can say anything about reaction rates...they all proceed too fast.

# Fig. 3C - Interphenazine ET with oxygen

Now let's setup the filenames, data columns etc, and we can go ahead and import and view the dataframe:

```{r}

#divide swv into rep and subrep and then subtract 1 from rep to match with GC
paths <-  dir(path='../../../data/Spectroscopy/phz_ET_vs_o2/', pattern = ".+[csv]$",recursive = T,full.names = T)

filenames <- basename(paths)

data_cols <-  c('t','blank','abs690')

filename_cols = c('year','month','day','PHZox','ox','PHZred','red','rep')

skip_rows=48 # or 48
n_max = 61
  

df <- import_to_df(filenames = filenames, 
                                       file_paths = paths, 
                                       data_cols = data_cols, 
                                       skip_rows = skip_rows, n_max = n_max,
                                       filename_cols = filename_cols,
                                       rep = T, PHZadded = F) %>% 
  mutate(phz_redox = paste(PHZox, ox, ' + ', PHZred, red, sep = ''))

df %>% kable() %>% kable_styling() %>% scroll_box(height = '300px')

```
<br>

Let's take a look at the conditions (triplicates) individually.

```{r}

ggplot(df, aes(x = t, y = abs690)) + geom_point(shape = 21) + facet_wrap(~phz_redox)

```

And now we can assemble the final plot:

```{r}

plot_phzET_o2 <- ggplot(df %>% filter(abs690>-0.01), aes(x = t-15, y = abs690, color = phz_redox)) + 
  geom_point(data = . %>% filter(!(phz_redox %in% c('PBSox + PCAred', 'PBSox + PCNred'))),
             shape = 21, alpha = 0.5, size = 0.5) + 
  geom_smooth(data = . %>% filter(abs690>0.01 & abs690<0.4) %>% 
                filter(!(phz_redox %in% c('PBSox + PCAred', 'PBSox + PCNred'))), 
              size = 0.5) + 
  geom_line(data = . %>% filter(phz_redox %in% c('PBSox + PCAred', 'PBSox + PCNred') & t >11), 
            aes(group = phz_redox), color = 'light gray', linetype = 2)+
  scale_color_manual(values = c('gray', "#56B4E9","#E69F00", "#000000")) +  
  labs(x = 'Time (sec)',y = 'Absorbance at 690nm') +  
  theme(legend.position = c(0.5, 0.8))

plot_phzET_o2
```


------

# Fig. 3E - DNA modified electrode (anoxic)

This data is electrochemical and contains the third/final sweep of an electrode with two different DNA monolayers.

```{r}
df_noO2 <- read_csv('../../../data/Electrochemistry/DNA_modified_electrode_noO2.csv',comment = "#") %>% 
  mutate(Construct=fct_relevel(Construct,'wm','mm'))

df_noO2 %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')
```

Let's make the figure:
```{r}

plot_dnaCT_noO2 <- ggplot(df_noO2,aes(x=E,y=Current, color = Construct))+
  geom_path()

plot_dnaCT_noO2_styled <- plot_dnaCT_noO2 +
  scale_x_reverse(labels = mV_label)+
  scale_y_continuous(labels = nA_label)+
  scale_color_manual(breaks = c('wm','mm'), labels=c("Well matched DNA","Mismatched DNA"), values = c("#66CCFF","#FFCC66")) +
  labs(x='E (mV vs. Ag/AgCl)',y='Current (nA)')+
  theme(legend.position = c(0.3, 0.8),
        legend.title = element_blank(),
        legend.background = element_rect(fill=NA))

plot_dnaCT_noO2_styled
```


# Fig. 3F - DNA modified electrode (with oxygen)

Let's read in the echem data:

```{r}
df_wO2 <- read_csv('../../../data/Electrochemistry/DNA_modified_electrode_wO2.csv',comment = "#") %>% 
  mutate(Condition=paste(Construct,Aerobic,PCN,sep='_')) %>% 
  mutate(Condition=fct_relevel(Condition,"wm_TRUE_TRUE","mm_TRUE_TRUE","wm_TRUE_FALSE"))

df_wO2 %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')
```

And let's make the figure:

```{r}
plot_dnaCT_wO2 <- ggplot(df_wO2 %>% filter(Aerobic==T),aes(x = E, y = Current, color = Condition))+
  geom_path()

plot_dnaCT_wO2_styled <- plot_dnaCT_wO2 +
  scale_x_reverse(labels = mV_label)+
  scale_y_continuous(labels = nA_label)+
  scale_color_manual(
    breaks = c('wm_TRUE_TRUE','mm_TRUE_TRUE','wm_TRUE_FALSE'), values = c("#66CCFF","#FFCC66","gray"),
    labels=c(expression("wmDNA"+"PCN"+O[2]),expression("mmDNA"+"PCN"+O[2]),expression("wmDNA"-"PCN"+O[2]))
    ) +
  labs(x='E (mV vs. Ag/AgCl)',y='Current (nA)')+
  theme(
    legend.position = c(0.3, 0.8),
    legend.title = element_blank(),
    legend.background = element_rect(fill=NA))

plot_dnaCT_wO2_styled
```

# Fig. 3H - Time resolved spectroscopy - raw data: liquid cells vs. biofilm

First let's read in the raw datasets. This uses the function `import_to_df()` from the `text_file_import.R` tool.

```{r}
data_path = '../../../data/Spectroscopy/Ru_Rh_time_resolved_spectroscopy/'

data_cols <-  c('time','intensity')

skip_rows=16

# Add 'reactor' to file name so it is parsed into column
filename_cols = c('date','rep','compressed')

data_names <- dir(path=data_path, pattern = ".+[dat]$")

# Add correct paths separate from filenames
data_paths <- data_names %>% paste(data_path, ., sep='')

# Read in all SWVs with one function call 
spec_data <- import_to_df(filenames = data_names, file_paths = data_paths, data_cols = data_cols, 
                          skip_rows = skip_rows, filename_cols = filename_cols, rep = T, PHZadded = F, n_max = 500)

spec_data %>% head() %>% kable(digits = 10) %>% kable_styling(bootstrap_options = 'condensed')

```

Next let's read in the metadata for the files. 

```{r}

metadata <- read_csv("../../../data/Spectroscopy/Ru_Rh_time_resolved_spectroscopy/time_resolved_spectroscopy_metadata.csv") %>% 
  group_by(material, material_id) %>% 
  mutate(run = rep - min(rep) + 1)


metadata %>% kable() %>% kable_styling(bootstrap_options = 'condensed') %>%
    scroll_box(width = "100%", height = "400px")

```

We will join the metadata with the raw data and make the plot comparing biofilm to liquid cells. We will only examine the first 6 conditions, since at higher quencher levels the signal exhibits complex behavior below the background signal.

```{r}
df_spec <- left_join(spec_data, metadata, by = 'rep') 

plot_raw_spec <- ggplot(df_spec %>% filter(run<= 6), 
                        aes(x = time, y = intensity, color = quencher_eq, group = quencher_eq)) + 
  geom_path() + 
  scale_x_continuous(limits = c(NA, 5e-7), labels = nA_label) + 
  labs(x = 'Time (ns)', y = 'Intensity (620nm)', color = 'Quencher\nEquivalents') + guides(color = F) +
  facet_wrap(~material, scale = 'free') 

plot_raw_spec


```

Now we will background subtract from the cell only acquisition - "run 1".

```{r}
df_spec_bg <- df_spec %>% 
  group_by(material, material_id, time) %>% 
  mutate(min = min(ifelse(run == 1, intensity, NA), na.rm = T)) %>% 
  mutate(bg_sub = intensity - min)

plot_bg_spec <- ggplot(df_spec_bg %>% filter(run <= 6 & bg_sub>=0), 
                       aes(x = time, y = bg_sub, color = quencher_eq, group = quencher_eq)) + 
  geom_path() + 
  scale_x_continuous(limits = c(NA, 5e-7), labels = nA_label) + 
  labs(x = 'Time (ns)', y = 'Intensity (620nm)', color = 'Quencher\nEquivalents') + 
  guides(color = F) + facet_wrap(~material)

plot_bg_spec

```

Clearly the biofilm signal shows the lifetime of the Ru probe, while the liquid sample shows mostly noise. There may be a small amount of signal in the absence of the quencher.

Now let's fit biexponential decays to the background subtracted biofilm datasets. This makes use of the self starting non-linear least squares biexponential function, `SSbiexp()`. This function is parameterized in the following way:

`output = A1*exp(-exp(lrc1)*input) + A2*exp(-exp(lrc2)*input)`

Where input is our time vector and output is the background subtracted intensity vector. A is the multiplier of the first (A1) and second (A2) exponential. Lrc is the natural log of the rate constant for the first (lrc1) and second (lrc2) exponential.

For figure S4, lrc parameters are converted into tau half lives `tau = 1/exp(lrc)`.

```{r}
df_spec_fit <- df_spec_bg %>% 
  filter(material == 'biofilm') %>% 
  filter(run %in% 2:6) %>% 
  filter(time>30e-9) %>% 
  filter(bg_sub>0) %>% 
  group_by(run, quencher_eq) %>% 
  nest()

fit_biexp_bg_sub <- function(df){
  
  A1 = 0.02; lrc1 = 16.5; A2 = 0.02; lrc2 = 15.7
  
  mod <- nls(bg_sub~SSbiexp(time, A1, lrc1, A2, lrc2), data = df)
  
  mod
}

df_spec_models <- df_spec_fit %>% 
  mutate(models = map(data, fit_biexp_bg_sub)) 

df_spec_preds <- df_spec_models %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)

df_spec_ests <- df_spec_models %>% 
  mutate(ests = map(models, tidy)) %>% 
  unnest(ests)


df_spec_ests %>% head()

```

Let's save these parameter estimates for use in supplemental figure S4. 

```{r}

write_csv(df_spec_ests %>% select(-data, -models), "phz2019_dna_ct_spec_biexp_fits.csv")

df_spec_ests %>% select(-data, -models) %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')

```

Now let's plot the background subtracted biofilm data with the model predictions.

```{r}

plot_bg_spec <- ggplot(df_spec_bg %>% filter(run %in% 1:6) %>% filter(material == 'biofilm'), aes(x = time, group = run, color = quencher_eq)) + 
  geom_point( aes(y = bg_sub),alpha = 0.2, shape = 21, size = 0.5) + 
  geom_path(data = df_spec_preds,aes(y = pred)) + scale_x_continuous(limits = c(NA,1e-6), labels = nA_label, name = 'Time (ns)') + 
  ylim(-0.001, NA) + labs(y = 'Background subtracted - Intensity (620nm)', color = 'Quencher\nequivalents')+
  guides(color = guide_colorbar(barwidth = 0.5, barheight = 2.5)) + theme(legend.position = c(0.75,0.75)) 

plot_bg_spec

```

# Create figure

```{r}

theme_set(theme_figure())


fig_3 <- plot_grid(plot_ctDNA_690, plot_ctDNA_690, plot_phzET_o2,
                   plot_dnaCT_noO2_styled, plot_dnaCT_noO2_styled, plot_dnaCT_wO2_styled,
          plot_raw_spec, plot_raw_spec, plot_bg_spec, 
          ncol = 3, scale = 0.95, align = 'hv', axis = 'tlr', labels = 'AUTO', label_size = 12)

fig_3

save_plot("../../../figures/phz2019_Fig_3.pdf", fig_3, base_height = 6, base_width = 7)
```

-----

```{r}
sessionInfo()
```