---
title: "Figure S7"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez-Otero, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

To see how we got from the raw electrochemical scans to the datasets used here, please see the following notebooks:

* [IDA blank processing](https://scott-saunders.github.io/phz_eDNA_2019/code/processing/IDA_blank/IDA_blank_processing.html)

This supplemental figure and notebook underlies some of the data in main figure 6, particularly the final panel. Specifically the model coefficients for the blank $D_{ap}$ Fig. S7C. These data are saved as .csv files in the directory containing this notebook.

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(broom)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../../tools/plotting_tools.R")


#Modify the plot theme
theme_set(theme_notebook())
```


# Fig. S7A

First we will import the blank IDA data that will be used for panels A-D. This data was processed from [this notebook](https://scott-saunders.github.io/phz_eDNA_2019/code/processing/IDA_blank/IDA_blank_processing.html), and includes SWV and GC scans taken with a blank IDA in solutions of known concentrations of PYO.

Let's read in the data and convert the µM units to $mol / cm^3$, which will be important to calculate $D_{ap}$ in correct units.

```{r}
df_soak <- read_csv("../../../processing/processed_data/phz_eDNA_2019_swv_gc_soak_blank.csv") %>% 
  mutate(PHZaddedInt = case_when(
    PHZadded == '100uM' ~ 100,
    PHZadded == '10uM' ~ 10,
    PHZadded == '25uM' ~ 25,
    PHZadded == '50uM' ~ 50,
    PHZadded == '75uM' ~ 75
  )) %>% 
  mutate(C_mol_cm3 = PHZaddedInt*(10^-9))
```

Here is the relationship between GC peak current and concentration:

```{r}

plot_GC <- ggplot(df_soak, aes(x = C_mol_cm3, y = signal_GC)) +    
  geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_GC_styled <- plot_GC+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(PYO~(mu*M)), y = expression(I[gc]~(nA)))
  
plot_GC_styled

```

$I_{GC}$ is simply $I = nFSDC$, so $D_{ap}$ is:

$$D_{ap} = \frac{m}{n F S}$$

We can define a simple function to calculate this $D_{ap}$ from the slope of the line in the above plot.

```{r}
dap_from_GC <- function(m){
  
  n=1.8
  S <-  18.4 #cm
  F_C = 96485.332
  
  d_ap <- m / (n * F_C * S)
  
  d_ap
}
```

Fitting the linear model and calculating $D_{ap}$ gives these values: 

```{r}
lm_GC <- tidy(lm(signal_GC~C_mol_cm3, data = df_soak), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_GC(m = estimate)) %>% 
  mutate(dap_high = dap_from_GC(m = conf.high)) %>% 
  mutate(dap_low = dap_from_GC(m = conf.low)) %>% 
  mutate(dataset = 'GC')

lm_GC %>% 
  kable() %>% 
  kable_styling()
```

<br>

# Fig. S7B

Now we will look at the relationship between concentration and peak SWV current.

```{r}
plot_SWV <- ggplot(df_soak, aes(x = C_mol_cm3, y = signal_SWV)) +
    geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_SWV_styled <- plot_SWV+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(PYO~(mu*M)), y = expression(I[swv]~(nA)))
  
plot_SWV_styled
```

For a plot of $I_{SWV}$ vs. $C$, the slope can be defined in terms of the parameters for $I_{SWV}$ other than concentration. This yields an expression for $D_{ap}$:

$$D_{ap} = \pi t_p \left( \frac{m}{\psi n F A} \right) ^2 $$

So we'll again define the function to calculate it:

```{r}
dap_from_SWV <- function(m, t_p=1/(2*300)){
  
  n=1.8
  psi <-  0.7
  A <-  0.025 #cm^2
  F_C = 96485.332
  
  d_ap <- pi * t_p * (m / (psi * n * F_C * A))^2
  
  d_ap
}
```

The linear model of this data gives the following:

```{r}
lm_SWV <- tidy(lm(signal_SWV~C_mol_cm3, data = df_soak), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_SWV(m = estimate)) %>% 
  mutate(dap_high = dap_from_SWV(m = conf.high)) %>% 
  mutate(dap_low = dap_from_SWV(m = conf.low)) %>% 
  mutate(dataset = 'SWV')

#glance(lm(signal_from_swv~C_mol_cm3, data = df_cal))

lm_SWV %>% 
  kable() %>% 
  kable_styling()
```

<br>

# Fig. S7C

Finally, we will examine the peak SWV current vs. the peak GC current, which is not dependent on the known PYO concentrations.

```{r}
plot_swvGC <- ggplot(df_soak, aes(x = signal_SWV, y = signal_GC)) + 
  geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_swvGC_styled <- plot_swvGC+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(I[swv]~(nA)), y = expression(I[gc]~(nA)))
  
plot_swvGC_styled
```

Recall that for this type of data:

$$D_{ap} = \frac{1}{\pi t_p} \left( \frac{m A \psi}{S} \right) ^2$$  

So, here's a function that will calculate $D_{ap}$ from the slope of this data:

```{r}
dap_from_swvGC <- function(m, t_p=1/(2*300)){
  
  psi <-  0.7
  A <-  0.025 #cm^2
  S <-  18.4 #cm
  
  d_ap <- (m*A*psi)^2 / (S^2 * pi * t_p)
  
  d_ap
}
```

And now, we'll fit a linear model and calculate a $D_{ap}$. We'll go ahead and save a csv of this result for use in main figure 6F.

```{r}
output_lm_swv_gc <- tidy(lm(signal_GC~signal_SWV, data = df_soak), conf.int = T) %>% 
  mutate(exp_id = 'Blank 1', run_id = 'Rep 1')

write_csv(output_lm_swv_gc, "phz2019_blank_Dap_lm_coefs.csv")

lm_swvGC <- tidy(lm(signal_GC~signal_SWV, data = df_soak), conf.int = T) %>% 
  filter(term=='signal_SWV') %>% 
  mutate(dap=dap_from_swvGC(m = estimate)) %>% 
  mutate(dap_high = dap_from_swvGC(m = conf.high)) %>% 
  mutate(dap_low = dap_from_swvGC(m = conf.low)) %>% 
  mutate(dataset = 'SWVvsGC')

lm_swvGC %>% 
  kable() %>% 
  kable_styling()
```

<br>

# Fig. S7D

Now let's compare the estimates from each subset of the data:

```{r}

lms <- bind_rows(lm_swvGC,lm_SWV,lm_GC) %>% 
  select(dataset, term, dap, dap_low, dap_high)

plot_cal_dap <- ggplot(lms, aes(x = dataset, y = dap)) + 
  geom_pointrange(aes(ymin = dap_low, ymax = dap_high)) + 
  ylim(0, 1e-5)

plot_cal_dap_styled <- plot_cal_dap+
  labs(x = 'Dataset', y = expression(D[ap]~(cm^2 / sec)))+
  scale_x_discrete(breaks = c('GC','SWV', 'SWVvsGC'), labels=c('GC vs.\n[PYO]', 'SWV vs.\n[PYO]', 'GC vs.\nSWV'))

plot_cal_dap_styled
```

# Fig. S7E

Now let's plot the SWV peak current decays for the blank IDA. This data comes from the blank IDA processing notebook.

```{r}

df_blank_swv <- read_csv("../../../processing/processed_data/phz_eDNA_2019_swv_blank_tran_time_signals.csv") %>% 
  filter(reactor == 'transfer' & PHZadded != '10uM') %>% 
  mutate(PHZadded = fct_relevel(PHZadded, c('25uM','50uM','75uM','100uM') ))

ggplot(df_blank_swv, aes(x = time, y = signal)) + 
  geom_point(shape = 21) + facet_wrap(~PHZadded, scales = 'free')
```

Now we will fit each of these decays. We will go ahead and save these coefficients as a csv, so that we can use them to calculate Dphys values in main figure 6.

```{r}
blank_nls <- df_blank_swv%>%
  group_by(PHZadded) %>% 
  do(
    tidy(
      nls(data = ., formula =  signal ~ b * (time)^-0.5 + a, start = c(b = 0.1, a = 1e-07) ),
      conf.int = T
    )
  ) %>% 
  arrange(desc(term))

# write results to csv for fig 6.
write_csv(blank_nls, "phz2019_blank_Dphys_nls_coefs.csv")

# print results here
blank_nls %>% kable(digits = 10) %>% kable_styling() %>% scroll_box(height = '300px')

```
<br>

Now we are going to take those coefficient estimates and predict datapoints that those values would generate at 1000 timepoints in the window our data is in. Essentially, we are going to generate the best fit line from those parameters and the 95% confidence interval so that we can plot it with the original datapoints.

```{r}
blank_grid <- tibble(time = seq(0.05, max(df_blank_swv$time), length.out = 1000))

blank_coefs <- left_join(blank_nls %>% filter(term == 'b'),
                      blank_nls %>% filter(term == 'a'), 
                      by = c('PHZadded'), suffix = c('_b','_a'))

blank_grid <- left_join(df_blank_swv %>% group_by(reactor, PHZadded) %>% summarise(),
                       blank_grid %>% mutate(reactor = 'transfer'), by = c('reactor'))

blank_grid_coef <- left_join(blank_grid, blank_coefs, by  = c('PHZadded'))

blank_preds <- blank_grid_coef %>% 
  mutate(pred = estimate_b * (time^-0.5) + estimate_a) %>% 
  mutate(pred_low = conf.low_b* (time^-0.5) + conf.low_a) %>% 
  mutate(pred_high = conf.high_b* (time^-0.5) + conf.high_a) 

write_csv(blank_preds, "phz2019_blank_Dphys_preds.csv")

# print simple results here
blank_preds %>% 
  select(reactor,PHZadded, time, pred, pred_high, pred_low, estimate_b, 
         conf.low_b, conf.high_b, estimate_a, conf.low_a, conf.high_a) %>% 
  head() %>% kable(digits = 7) %>% kable_styling() %>% scroll_box(height = '300px')
```
<br>

Now they we have those predictions we can plot everything together:

```{r}
plot_decays_blank <- ggplot(blank_preds, aes(x = time, y = pred)) + 
  geom_ribbon(aes(ymin = pred_low, ymax = pred_high), fill = 'light gray') +
  geom_path(linetype = 2)+
  geom_point(data =df_blank_swv, aes(x = time, y = signal) , shape = 21)+
  facet_wrap(~PHZadded, scale = 'free')

plot_decays_blank_styled <- plot_decays_blank+
    scale_y_continuous(labels = nA_label) +  
  labs(y = expression(I[swv]~(nA)), x = 'Time (min)', title = 'Blank')

plot_decays_blank_styled
```


# Create figure

```{r}
theme_set(theme_figure())


fig_s7_top <- plot_grid(plot_GC_styled, plot_SWV_styled, 
          plot_swvGC_styled, plot_cal_dap_styled, 
          align = 'hv', axis = 'tblr',scale = 0.95, labels = 'AUTO', label_size = 12)

fig_s7_bottom <- plot_grid(plot_decays_blank_styled, 
          align = 'hv', axis = 'tblr',scale = 0.95, labels = 'E', label_size = 12)

fig_s7 <- plot_grid(fig_s7_top, fig_s7_bottom, ncol = 1, rel_heights = c(1.25,1))

fig_s7

save_plot("../../../../figures/supplement/phz2019_Fig_S7.pdf", fig_s7, base_height = 6, base_width = 5)
```

-----

```{r}

sessionInfo()

```
