---
title: "Figure S7"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez-Otero, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

To see how we got from the raw electrochemical scans to the datasets used here, please see the following notebooks:

* [IDA blank processing](https://scott-saunders.github.io/phz_eDNA_2019/code/processing/IDA_blank/IDA_blank_processing.html)

This supplemental figure and notebook underlies some of the data in main figure 6, particularly the final panel. Specifically the model coefficients for the blank $D_{ap}$ Fig. S7C. These data are saved as .csv files in the directory containing this notebook.

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(broom)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../../tools/plotting_tools.R")


#Modify the plot theme
theme_set(theme_notebook())
```


# Fig. S7A

First we will import the blank IDA data that will be used for panels A-D. This data was processed from [this notebook](https://scott-saunders.github.io/phz_eDNA_2019/code/processing/IDA_blank/IDA_blank_processing.html), and includes SWV and GC scans taken with a blank IDA in solutions of known concentrations of PYO.

Let's read in the data and convert the µM units to $mol / cm^3$, which will be important to calculate $D_{ap}$ in correct units.

```{r}
df_soak <- read_csv("../../../processing/processed_data/phz_eDNA_2019_swv_gc_soak_blank.csv") %>% 
  mutate(PHZaddedInt = case_when(
    PHZadded == '100uM' ~ 100,
    PHZadded == '10uM' ~ 10,
    PHZadded == '25uM' ~ 25,
    PHZadded == '50uM' ~ 50,
    PHZadded == '75uM' ~ 75
  )) %>% 
  mutate(C_mol_cm3 = PHZaddedInt*(10^-9))
```

Here is the relationship between GC peak current and concentration:

```{r}

plot_GC <- ggplot(df_soak, aes(x = C_mol_cm3, y = signal_GC)) +    
  geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_GC_styled <- plot_GC+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(PYO~(mu*M)), y = expression(I[gc]~(nA)))
  
plot_GC_styled

```

$I_{GC}$ is simply $I = nFSDC$, so $D_{ap}$ is:

$$D_{ap} = \frac{m}{n F S}$$

We can define a simple function to calculate this $D_{ap}$ from the slope of the line in the above plot.

```{r}
dap_from_GC <- function(m){
  
  n=1.8
  S <-  18.4 #cm
  F_C = 96485.332
  
  d_ap <- m / (n * F_C * S)
  
  d_ap
}
```

Fitting the linear model and calculating $D_{ap}$ gives these values: 

```{r}
lm_GC <- tidy(lm(signal_GC~C_mol_cm3, data = df_soak), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_GC(m = estimate)) %>% 
  mutate(dap_high = dap_from_GC(m = conf.high)) %>% 
  mutate(dap_low = dap_from_GC(m = conf.low)) %>% 
  mutate(dataset = 'GC')

lm_GC %>% 
  kable() %>% 
  kable_styling()
```

<br>

# Fig. S7B

Now we will look at the relationship between concentration and peak SWV current.

```{r}
plot_SWV <- ggplot(df_soak, aes(x = C_mol_cm3, y = signal_SWV)) +
    geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_SWV_styled <- plot_SWV+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(PYO~(mu*M)), y = expression(I[swv]~(nA)))
  
plot_SWV_styled
```

For a plot of $I_{SWV}$ vs. $C$, the slope can be defined in terms of the parameters for $I_{SWV}$ other than concentration. This yields an expression for $D_{ap}$:

$$D_{ap} = \pi t_p \left( \frac{m}{\psi n F A} \right) ^2 $$

So we'll again define the function to calculate it:

```{r}
dap_from_SWV <- function(m, t_p=1/(2*300)){
  
  n=1.8
  psi <-  0.7
  A <-  0.025 #cm^2
  F_C = 96485.332
  
  d_ap <- pi * t_p * (m / (psi * n * F_C * A))^2
  
  d_ap
}
```

The linear model of this data gives the following:

```{r}
lm_SWV <- tidy(lm(signal_SWV~C_mol_cm3, data = df_soak), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_SWV(m = estimate)) %>% 
  mutate(dap_high = dap_from_SWV(m = conf.high)) %>% 
  mutate(dap_low = dap_from_SWV(m = conf.low)) %>% 
  mutate(dataset = 'SWV')

#glance(lm(signal_from_swv~C_mol_cm3, data = df_cal))

lm_SWV %>% 
  kable() %>% 
  kable_styling()
```

<br>

# Fig. S7C

Finally, we will examine the peak SWV current vs. the peak GC current, which is not dependent on the known PYO concentrations.

```{r}
plot_swvGC <- ggplot(df_soak, aes(x = signal_SWV, y = signal_GC)) + 
  geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_swvGC_styled <- plot_swvGC+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(I[swv]~(nA)), y = expression(I[gc]~(nA)))
  
plot_swvGC_styled
```

Recall that for this type of data:

$$D_{ap} = \frac{1}{\pi t_p} \left( \frac{m A \psi}{S} \right) ^2$$  

So, here's a function that will calculate $D_{ap}$ from the slope of this data:

```{r}
dap_from_swvGC <- function(m, t_p=1/(2*300)){
  
  psi <-  0.7
  A <-  0.025 #cm^2
  S <-  18.4 #cm
  
  d_ap <- (m*A*psi)^2 / (S^2 * pi * t_p)
  
  d_ap
}
```

And now, we'll fit a linear model and calculate a $D_{ap}$. We'll go ahead and save a csv of this result for use in main figure 6F.

```{r}
output_lm_swv_gc <- tidy(lm(signal_GC~signal_SWV, data = df_soak), conf.int = T) %>% 
  mutate(exp_id = 'Blank 1', run_id = 'Rep 1')

write_csv(output_lm_swv_gc, "phz2019_blank_Dap_lm_coefs.csv")

lm_swvGC <- tidy(lm(signal_GC~signal_SWV, data = df_soak), conf.int = T) %>% 
  filter(term=='signal_SWV') %>% 
  mutate(dap=dap_from_swvGC(m = estimate)) %>% 
  mutate(dap_high = dap_from_swvGC(m = conf.high)) %>% 
  mutate(dap_low = dap_from_swvGC(m = conf.low)) %>% 
  mutate(dataset = 'SWVvsGC')

lm_swvGC %>% 
  kable() %>% 
  kable_styling()
```

<br>

# Fig. S7D

Now let's compare the estimates from each subset of the data:

```{r}

lms <- bind_rows(lm_swvGC,lm_SWV,lm_GC) %>% 
  select(dataset, term, dap, dap_low, dap_high)

plot_cal_dap <- ggplot(lms, aes(x = dataset, y = dap)) + 
  geom_pointrange(aes(ymin = dap_low, ymax = dap_high)) + 
  ylim(0, 1e-5)

plot_cal_dap_styled <- plot_cal_dap+
  labs(x = 'Dataset', y = expression(D[ap]~(cm^2 / sec)))+
  scale_x_discrete(breaks = c('GC','SWV', 'SWVvsGC'), labels=c('GC vs.\n[PYO]', 'SWV vs.\n[PYO]', 'GC vs.\nSWV'))

plot_cal_dap_styled
```

# Create figure

```{r}
theme_set(theme_figure())


fig_s7 <- plot_grid(plot_GC_styled, plot_SWV_styled, 
          plot_swvGC_styled, plot_cal_dap_styled, 
          align = 'hv', axis = 'tblr',scale = 0.95, labels = 'AUTO', label_size = 12)

fig_s7

save_plot("../../../../figures/supplement/phz2019_Fig_S7.pdf", fig_s7, base_height = 4, base_width = 5)
```

-----

```{r}

sessionInfo()

```
