---
title: "Figure 2"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez Otero, Scott A. Trammell, Eric D.A. Stemp, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

Panel A of figure 2 is isothermal titration calorimetry data that was analyzed outside of R. 

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Check packages
source("../../tools/package_setup.R")

# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../tools/plotting_tools.R")

# Modify the plot theme 
theme_set(theme_notebook())
```

# Fig. 2B - WT Colony +/- DNase

Let's read in the dnase dataset and look at the table:
```{r}
dnase_extracts <- read_csv('../../../data/LC-MS/2018_10_08_HPLC_concentrations_df.csv',comment = "#") %>% 
  filter(Strain=='WT' & Day=='D4') %>% 
  group_by(Phenazine,Condition,Material) %>% 
  mutate(mean = ifelse(Replicate==1, mean(calcConc), NA))
  

dnase_extracts %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')

```

Now, let's look at an overview of the experiment.

```{r}

ggplot(dnase_extracts ,aes(x=Condition,y=calcConc))+
  geom_col( aes(y = mean), fill = 'light gray') +
    geom_jitter(width=0.1,height=0,shape=21,size=1)+
    facet_wrap(Material~Phenazine,scales='free')

```

And let's do a statistical test to compare the DNase +/- treatments. 

First, a t-test for whether or not agar with DNase concentrations were higher than the 'none' treatment:

```{r}

dnase_extracts %>% 
  spread(Condition,calcConc) %>% 
  filter(Material=='agar') %>% 
  group_by(Material,Phenazine) %>% 
  summarise(conf_int_low = t.test( DNase,none, alternative = 'greater')$conf.int[1],
            conf_int_high = t.test( DNase,none, alternative = 'greater')$conf.int[2],
            p_value = t.test( DNase,none, alternative = 'greater')$p.value)
```
There are no significant differences p<0.05.

Second, a t-test for whether or not biofilm (aka cell) with DNase concentrations were lower than the 'none' treatment
```{r}
  
dnase_extracts %>% 
  spread(Condition,calcConc) %>% 
  filter(Material=='cells') %>% 
  group_by(Material,Phenazine) %>% 
  summarise(conf_int_low = t.test( DNase, none, alternative = 'less')$conf.int[1],
            conf_int_high = t.test(DNase, none,  alternative = 'less')$conf.int[2],
            p_value = t.test( DNase, none, alternative = 'less')$p.value)
```
There is a significant difference for PYO and PCN.

Here you can see that the agar concentrations between the Dnase treated and untreated don't differ meaningfully, but the cell/biofilm concentrations might. This might be because for this experiment the colonies were transferred to a fresh agar plate for only 24hrs as opposed to staying on the same plate for 4 days as with the pel experiment. So, let's ignore the agar concentrations for now. It's also important to note that by calculating a ratio as we did above for pel we do risk amplifying meaningless differences by dividing large numbers by small numbers.

Here's the biofilm only:

```{r}
# Plot layout
dnase_plot <- ggplot(dnase_extracts %>% filter(Material=='cells') ,aes(x=Condition,y=calcConc))+
  geom_col( aes(y = mean), fill = 'light gray') +
    geom_jitter(width=0.1,height=0,shape=21,size=1)+
    facet_wrap(~Phenazine,scales='free') 

# Plot styling
dnase_plot_styled <- dnase_plot +
  labs(x=NULL, y=expression("Biofilm concentration" ~ (mu*M ))) + 
  scale_x_discrete(limits = c("none",'DNase')) +
  guides(fill = F)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

dnase_plot_styled

```

# Fig. 2C - ∆pel colony

Let's read in and look at the table of pel data. 
```{r}
pel_extracts <- read_csv('../../../data/LC-MS/2018_10_30_HPLC_concentrations_df.csv',comment = "#") %>% 
  filter(strain %in% c('WTpar','dPel'))

pel_extracts %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')
```

Ok, now let's plot an overview of the dataset. 
```{r}
pel_extracts_means <- pel_extracts %>% 
  group_by(material, measured_phenazine, strain) %>% 
  mutate(mean = ifelse(replicate==1, mean(calcConc), NA)) 

ggplot(pel_extracts_means ,aes(x=strain,y=calcConc))+
  geom_col( aes(y = mean), fill = 'light gray') +
    geom_jitter(width=0.1,height=0,shape=21,size=2)+
    facet_wrap(material~measured_phenazine,scales='free')+
    scale_x_discrete(breaks = c('dPel','WTpar'), labels=c(expression(Delta*"pel"),"WT")) +
    labs(x='Strain',y=expression("Biofilm concentration" ~ (mu*M )) ) +
    guides(fill=F)
```
You can see that for each phenazine the concentration differs between the strains for both the cells aka biofilm and the agar.

And let's perform the t-test on whether or not ∆pel concentrations are different than WT concentrations. 

```{r}
pel_extracts %>% 
  spread(strain,calcConc) %>% 
  group_by( material ,measured_phenazine) %>% 
  summarise(conf_int_low = t.test(dPel, WTpar,  alternative = 'two.sided')$conf.int[1],
            conf_int_high = t.test(dPel,WTpar, alternative = 'two.sided')$conf.int[2],
            p_value = t.test( dPel,WTpar, alternative = 'two.sided')$p.value)
```

Both the agar and biofilm concentrations for PYO and PCN are statistically significantly different in ∆pel vs. WT strains.

Let's calculate the retention ratio.

```{r}

# Split dataset by material
pel_extracts_means_agar <- pel_extracts_means %>% 
  filter(material=='agar')

pel_extracts_means_biofilm <- pel_extracts_means %>% 
  filter(material=='cells')

# Join agar and cell observations and calculate retention ratios = biofilm / agar
pel_extracts_means_join <- left_join(pel_extracts_means_biofilm, 
                                     pel_extracts_means_agar, 
                                     by=c('strain','replicate','measured_phenazine'), 
                                     suffix = c('_from_biofilm','_from_agar') 
                                     ) %>% 
  mutate(retention_ratio = calcConc_from_biofilm / calcConc_from_agar) %>% 
  mutate(mean_retention_ratio = mean_from_biofilm / mean_from_agar)

# Plot Layout
plot_pel <- ggplot(pel_extracts_means_join ,aes(x=strain,y=retention_ratio))+
  geom_col( aes(y = mean_retention_ratio), fill = 'light gray') +
    geom_jitter(width=0.1,height=0,shape=21,size=1)+
    facet_wrap(~measured_phenazine,scales='free')

# Styling
plot_pel_styled <- plot_pel +
  scale_x_discrete(breaks = c('WTpar','dPel'), 
                   labels=c("WT",expression(Delta*"pel")), 
                   limits = c('WTpar','dPel')) +
  scale_y_continuous(labels = fold_label)+
  labs(x=NULL, y = '[Biofilm] / [Agar]') +
  guides(fill=F)  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_pel_styled

```

And let's perform the t-test on whether or not ∆pel ratios are greater than WT ratios. 

```{r}
pel_extracts_means_join %>% 
  spread(strain,retention_ratio) %>% 
  group_by(measured_phenazine) %>% 
  summarise(conf_int_low = t.test(dPel, WTpar,  alternative = 'greater')$conf.int[1],
            conf_int_high = t.test(dPel,WTpar, alternative = 'greater')$conf.int[2],
            p_value = t.test( dPel,WTpar, alternative = 'greater')$p.value)

```

There is a statistically significant difference for PYO. 

# Fig. 2D - EtBr vs. PHZ in colonies

Let's read in the data and calculate the concentrations.
```{r}
df_dphz_2 <- read_csv("../../../data/LC-MS/2019_07_23_colony_HPLC_dPHZ_data_2.csv")

df_dphz_2 <- df_dphz_2 %>% 
  mutate(condition_conc = fct_relevel(condition_conc, c('0uM','10uM','100uM','200uM','500uM'))) %>% 
    mutate(condition = fct_relevel(condition, c('PBS','etbr','dmso','pi')))

df_dphz_conc <- df_dphz_2 %>% 
  filter(added_phz == measured_phz) %>% 
  mutate(measured_phz_conc = case_when(
    material == 'cells' ~ (Amount * 2) * (800 / 60),
    material == 'agar' ~ (Amount * 2) * (8 / 5)
  )) %>% 
  group_by(measured_phz, strain, condition, condition_conc) %>% 
  mutate(mean_phz_conc = mean(measured_phz_conc))
  
```

And let's plot:

```{r}
plot_dphz_etbr <- ggplot(df_dphz_conc %>% filter(condition %in% c('etbr','PBS')), 
       aes(x = condition_conc, y = measured_phz_conc)) + 
  geom_col(data = df_dphz_conc %>% filter(condition %in% c('etbr','PBS') & rep ==1), 
           aes(x = condition_conc, y  = mean_phz_conc), fill = 'light gray') +
  geom_point(shape = 21, size = 1)  + facet_wrap(~measured_phz, scales = 'free') + ylim(0, NA) + 
  labs(x = expression('EtBr added to agar'~(mu*M)), y = expression('Biofilm phz concentration'~(mu*M)))+
  scale_x_discrete(labels = c(0, 10, 100, 500)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
plot_dphz_etbr
```


# Fig. 2E - WT eDNA with TOTO-1

Let's read in the standards, biofilm and metadata:
```{r}

df_stds <- read_csv("../../../data/Spectroscopy/2019_11_22_std_preCTdna.csv") %>%  gather(key = "well", value = "FluorInt", -wavelength) %>% mutate(read = 1)

df_biofilms <- read_csv("../../../data/Spectroscopy/2019_11_22_std_wt_dphz_postCTdna.csv") %>%  gather(key = "well", value = "FluorInt", -wavelength) %>% mutate(read = 2)

df_meta <- read_csv("../../../data/Spectroscopy/2019_11_22_well_metadata.csv")

df_toto <- left_join(bind_rows(df_stds, df_biofilms), df_meta, by = c('well')) %>% filter(wavelength == 535) %>% filter(!(strain == 'std' & read == 2))

df_toto %>% kable() %>% kable_styling(bootstrap_options = 'condensed') %>%
    scroll_box(width = "100%", height = "400px")
```

Now we can make the plot:
```{r}
std_levels <- df_toto %>% filter(strain == 'std') %>% filter(FluorInt <30000 & FluorInt >5000) %>% 
  mutate(eDNA_ug_mL = well_std_conc * 4 * 800 / 60) %>% 
  mutate(eDNA_dsDNA_uM = eDNA_ug_mL * 150 / 50) %>% 
  mutate(labels = paste(eDNA_dsDNA_uM, 'mu*M', sep = '~'))

bio_reps <- df_toto %>% filter(strain %in% c('WT','dPHZ') & ctDNA_added == FALSE & toto_added == T) %>% 
  group_by(strain, bio_rep) %>% 
  summarise(mean = mean(FluorInt), sd = sd(FluorInt))

plot_toto <- ggplot(bio_reps, aes(x = strain, y = mean)) +
  geom_hline(yintercept = std_levels$FluorInt, linetype = 2, color = 'light gray', label = labels)+
  geom_pointrange(aes(ymin = mean - sd, ymax = mean + sd), position = position_jitter(width = 0.15, height = 0), size = 0.1) + 
  geom_text(data = std_levels, aes(x = c(2.5,2.5,2.5,2.5), y = FluorInt - 1000, label = labels),parse = T, color = 'light gray', size = (6 * 5 / 14))+
  ylim(0,NA) 

plot_toto_styled <- plot_toto + 
  labs(x = NULL, y = 'TOTO-1 Fluorescence (A.U.)') + 
  scale_x_discrete(breaks = c('dPHZ','WT'), 
                   labels=c(expression(Delta*"phz*"), 'WT')) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text.y = element_text(angle = 45, hjust = 1))

plot_toto_styled
```

# Create Figure

Let's put everything together:
```{r}

theme_set(theme_figure())

top_panels <- plot_grid(dnase_plot_styled, plot_pel_styled, ncol = 2, labels = c('B','C'), scale = 0.95, 
                        label_size = 12, align = 'hv', axis = 'tblr')

bottom_panels <- plot_grid(plot_dphz_etbr, plot_toto_styled, ncol = 2, labels = c('D','E'), 
                           scale = 0.95, label_size = 12, align = 'hv', axis = 'lr')

all_right_panels <- plot_grid(top_panels, bottom_panels, ncol = 1, scale = 1.0, align = 'hv', axis = 'tblr')

fig_2 <- plot_grid(NULL, all_right_panels, ncol = 2, rel_widths = c(0.5, 1), labels = c('A',''), label_size = 12)

fig_2

save_plot("../../../figures/phz2019_Fig_2.pdf", fig_2, base_width = 7, base_height = 3.5)
```

------

```{r}

sessionInfo()

```