---
title: "Figure S1"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez Otero, Scott A. Trammell, Eric D.A. Stemp, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

Fig S1A and S1B are a sets of images.

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Check packages
source("../../../tools/package_setup.R")

# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(broom)
library(modelr)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../../tools/plotting_tools.R")


#Modify the plot theme
theme_set(theme_notebook())
```

# Fig. S1C - Biofilms without membrane

Let's read in the data and we will calculate the biofilm concentration by assuming the colony was about 200 µL in size and resuspended in 3 mL, while the agar is the same as elsewhere (5 mL volume with 3 mL PBS added = 8 mL total).

```{r}
df <- read_csv('../../../../data/LC-MS/WT_noMem_HPLC_03_08_18.csv')

#assumes colonies were ~200uL in size, resuspended in 3mL. 
df_corrected <- df %>% 
  mutate(phzConc=ifelse(Material=='biofilm',Amount * (3000 / 200), Amount * (8 / 5))) %>% 
  group_by(Material, Name) %>% 
  mutate(mean = ifelse(Replicate==1,mean(phzConc),NA))

noMem_plot <- ggplot(df_corrected,aes(x=Material,y=phzConc))+
  geom_col(aes(y = mean, fill = Material)) + 
  geom_jitter(height = 0, width = 0.1, shape = 21, size =1) +
  facet_wrap(~Name, scales = 'free')

#Plot styling
noMem_plot_styled <- noMem_plot +
  labs(x = NULL, y = expression("Phenazine concentration" ~ ( mu*M )), title = 'WT: no membrane') + 
  scale_fill_manual(guide = F, values = c("#66CCFF","#FFCC66")) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
noMem_plot_styled     
```

# Fig. S1D - Biofilms with sonication

Let's read in the +/- sonication data and calculate the concentration as done in Fig. 1:
```{r}
wtSon_pdaMan <- read_csv("../../../../data/LC-MS/WTsonication_PDAmanual_08_29_18.csv") %>% 
  mutate(calcConc = Amount*2*(800 / 62)) %>% #Dilution factor
  group_by(measured_phenazine, Condition) %>% 
  mutate(mean = ifelse(Rep==1,mean(calcConc),NA))

# Plot layout
wtSon_pdaMan_plot <- ggplot(wtSon_pdaMan, aes(x = Condition, y = calcConc)) +
  geom_col(aes(y = mean), fill = 'light gray')+
  geom_jitter(shape = 21, height = 1, width = 0.1, size =1) + 
  facet_wrap(~measured_phenazine, scales = 'free') + 
  ylim(0,NA)

#Plot styling
wtSon_pdaMan_plot_styled <- wtSon_pdaMan_plot +
  labs(x = 'Sonication', y = expression("Biofilm concentration" ~ ( mu*M )), title = 'WT: sonication') + 
  scale_x_discrete(breaks = c('noSon','withSon'), 
                   labels=c("-","+"))
    
wtSon_pdaMan_plot_styled     
```

It looks like there may be a slightly higher biofilm concentration upon sonication. Let's perform a statistical test:

```{r}
wtSon_pdaMan %>% 
  spread(Condition,calcConc) %>% 
  group_by(measured_phenazine) %>% 
  summarise(conf_int_low = t.test(withSon,noSon,   alternative = 'greater')$conf.int[1],
            conf_int_high = t.test(withSon, noSon,alternative = 'greater')$conf.int[2],
            p_value = t.test( withSon,noSon, alternative = 'greater')$p.value)
```

It appears there is a statistically significant difference for PCN at p<0.05.

# Fig. S1E - ∆phz* retention

Let's read in the ∆phz* retention data.
```{r}

pdaMan_data <- read_csv("../../../../data/LC-MS/dPHZstar_PDAmanual_PHZretention_08_29_18.csv") %>% 
  mutate(calcConc = Amount * 2 *(800 / 62) ) %>% 
  group_by(Condition,Day,measured_phenazine) %>% 
  mutate(mean = ifelse(Rep==1,mean(calcConc),NA))
```

Here's a plot of the data for all 3 phenazines incubated simultaneously:
```{r}
dphz_ret_pdaMan_plot <- ggplot(pdaMan_data %>% filter(Condition =='PHZ'), aes(x = Day, y = calcConc, )) + 
  geom_col(aes(y = mean,fill = Day))+
  geom_jitter(height = 0, width = 0.1, shape = 21) +
  facet_wrap(~measured_phenazine, scales = 'free') + 
  ylim(0,NA)

#Plot styling
dphz_ret_pdaMan_plot_styled <- dphz_ret_pdaMan_plot +
  labs(x = NULL, y = "Phenazine Concentration") + 
  theme(axis.title.x = element_text(size = 14)) + 
  scale_fill_manual(guide = F, values = c("#66CCFF","#FFCC66")) + 
  scale_x_discrete(breaks = c('D3','D4'), 
                   labels=c("Day 3","Day 4"))

dphz_ret_pdaMan_plot_styled

```

And here is the final plot where ∆phz* biofilms were incubated with individual phenazines:

```{r}

pdaMan_indPhz_plot <- ggplot(pdaMan_data %>% filter(Condition == measured_phenazine), aes(x = Day, y = calcConc, )) + 
  geom_col(aes(y = mean), fill='light gray')+
  geom_jitter(height = 0, width = 0.1, shape = 21, size = 1) +
  facet_wrap(~measured_phenazine, scales = 'free') + 
  ylim(0,NA)

#Plot styling
pdaMan_indPhz_plot_styled <- pdaMan_indPhz_plot +
  labs(x = NULL, y = expression("Biofilm concentration" ~ (mu*M )), title = '∆phz*') + 
  theme(axis.title.x = element_text(size = 14),axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(guide = F, values = c("#66CCFF","#FFCC66")) + 
  scale_x_discrete(breaks = c('D3','D4'), labels=c("Day 3","Day 4"))

pdaMan_indPhz_plot_styled
```


# Create figure

```{r}

theme_set(theme_figure())

fig_s1 <- plot_grid(noMem_plot_styled, wtSon_pdaMan_plot_styled, pdaMan_indPhz_plot_styled, 
                    labels = c('C','D','E'), ncol = 3, label_size = 12, scale = 0.95)

fig_s1

save_plot("../../../../figures/supplement/phz2019_Fig_S1.pdf", fig_s1,base_width = 7, base_height = 2)

```

-------

```{r}
sessionInfo()
```