---
title: "Figure S6"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez-Otero, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

To see how we got from the raw electrochemical scans to the datasets used here, please see the following notebooks:

* [IDA ∆phz biofilm processing](https://scott-saunders.github.io/phz_eDNA_2019/code/processing/IDA_dPHZ/IDA_dPHZ_processing.html)
* [IDA WT biofilm processing](https://scott-saunders.github.io/phz_eDNA_2019/code/processing/IDA_WT/IDA_WT_processing.html)
* [IDA blank processing](https://scott-saunders.github.io/phz_eDNA_2019/code/processing/IDA_blank/IDA_blank_processing.html)

This supplemental figure and notebook underlies some of the data in main figure 6, particularly the final panel. Specifically the model coefficients for $D_{ap}$ and $D_{phys}$ are calculated in the Fig. S6B-D sections. These data are saved as .csv files in the directory containing this notebook.


----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(broom)
library(modelr)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../../tools/plotting_tools.R")


#Modify the plot theme
theme_set(theme_notebook())
```

# Fig. S6A

Let's compare WT and ∆phz* + PYO biofilms. We'll go ahead and read in some of the processed data from the WT and ∆phz processing notebooks. First we will look at the decays of the SWV peak currents:

```{r}
df_WT <- read_csv("../../../processing/processed_data/phz_eDNA_2019_swv_gc_WT_signals.csv") %>% 
  mutate(strain = 'WT')

df_dphz_1 <- read_csv("../../../processing/processed_data/phz_eDNA_2019_swv_gc_signals.csv") %>% 
  filter(reactor == 'transfer' & exp == 1 & run == 1) %>%
  mutate(strain = 'dPHZ')

df_combined <- bind_rows(df_dphz_1, df_WT)


plot_decay_combo <- ggplot(df_combined, aes(x = time_SWV, y = signal_SWV, fill = strain, color = strain)) + 
  geom_line() + geom_point(shape = 21, color = 'black') 

plot_decay_combo_styled <- plot_decay_combo + 
  scale_y_continuous(labels = nA_label) +  
  labs(y = expression(I[swv]~(nA)), x = 'Time (min)')
  
plot_decay_combo_styled
```

Now let's plot the SWV vs. GC plots that we use to calculate $D_{ap}$:

```{r}
plot_dap_combo <- ggplot(df_combined %>% filter(rep>0), aes(x = signal_SWV, y = signal_GC, fill = strain)) + 
  geom_smooth(method = 'lm', color = 'black', fill = 'light gray',linetype = 2, aes(group = strain)) + 
  geom_point(shape = 21)  
plot_dap_combo_styled <- plot_dap_combo+
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(I[swv]~(nA)), y = expression(I[gc]~(nA)))

plot_dap_combo_styled
```

# Fig. S6B

First, we will read in the paired SWV and GC signals that were generated from the processing notebook. Then we can plot the SWV vs. GC signals for both biofilms and technical replicates. Each dataset is fit with a linear model shaded with a 95% confidence interval.

```{r}

df_swv_gc <- read_csv("../../../processing/processed_data/phz_eDNA_2019_swv_gc_signals.csv") %>% 
  mutate(exp_id = ifelse(exp =='2', 'Biofilm 1', 'Biofilm 2')) %>% 
  mutate(run_id = paste('Rep ',run, sep = ''))

plot_swv_gc_dphz <- ggplot(df_swv_gc %>% filter(reactor == 'transfer' & rep>0), 
       aes(x = signal_SWV, y = signal_GC)) + 
  geom_smooth(method='lm', linetype = 2, color = 'black')+
  geom_point(shape = 21) + 
  facet_wrap(exp_id~run_id,scales='free')+
  scale_color_viridis(discrete = T)

plot_swv_gc_dphz_styled <- plot_swv_gc_dphz + 
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(I[swv]~(nA)), y = expression(I[gc]~(nA)), title = '∆phz*')

plot_swv_gc_dphz_styled

```

Above we generated the linear models on the fly for plotting, but let's actually generate those models so that we can look at the quality of each fit. These R squared values will be added to each panel in illustrator.

```{r}
mods_swv_gc <- df_swv_gc %>% 
  filter(reactor == 'transfer' & rep>0) %>% 
  group_by(exp_id, run_id) %>% 
  do(fit = lm(signal_GC~signal_SWV,data = .))

mods_swv_gc_glance <- glance(mods_swv_gc, fit) %>% 
  select(exp_id,run_id,r.squared,adj.r.squared, p.value)

mods_swv_gc_glance %>% kable(digits = 20) %>%  kable_styling() %>% scroll_box(height = '300px')
```

<br>

Here are the actual coefficients from those models, including 95% confidence intervals. We will use these coefficients to calculate Dap in Fig. 6. We will write this dataframe to a csv for later use.

```{r}

mods_swv_gc_tidy <- tidy(mods_swv_gc, fit, conf.int=T) %>% arrange(desc(term))

write_csv(mods_swv_gc_tidy, "phz2019_dPHZ_Dap_lm_coefs.csv")

mods_swv_gc_tidy %>% kable() %>%  kable_styling() %>% scroll_box(height = '300px')
```

<br>

# Fig. S6C

First let's read in the data, which are the peak SWV current signals over time from the ∆phz* biofilms:

```{r}
df_dphz_swv <- read_csv("../../../processing/processed_data/phz_eDNA_2019_signals_long.csv") %>% 
  filter(echem == 'SWV') %>% 
  filter(electrode == 'i1' & reactor %in% c('transfer','soak')) %>% 
  mutate(exp_id = ifelse(exp =='2', 'Biofilm 1', 'Biofilm 2')) %>% 
  mutate(run_id = paste('Rep ',run, sep = ''))

ggplot(df_dphz_swv %>% filter(reactor == 'transfer'), aes(x = time, y = signal)) + 
  geom_point(shape = 21) + facet_wrap(exp_id~run_id, scales = 'free')
```

Now we will fit each of these decays with the expression: 
$$y = b (x)^{-0.5} + a$$ 

We will fit using a nonlinear least squares method, the `nls()` function. Here you can see the model coefficient estimates and confidence intervals for each data set. We will go ahead and save these coefficients as a csv, so that we can use them to calculate $D_{phys}$ values in main figure 6.

```{r}
dphz_nls <- df_dphz_swv %>% filter(reactor == 'transfer') %>%
  group_by(exp, run) %>% 
  do(
    tidy(
      nls(data = ., formula =  signal ~ b * (time)^-0.5 + a, start = c(b = 0.1, a = 1e-07) ),
      conf.int = T
    )
  ) %>% 
  arrange(desc(term))

# write results to csv for fig 6.
write_csv(dphz_nls, "phz2019_dPHZ_Dphys_nls_coefs.csv")

# print results here
dphz_nls %>% kable(digits = 10) %>% kable_styling() %>% scroll_box(height = '300px')

```
<br>

Now we are going to take those coefficient estimates and predict datapoints that those values would generate at 1000 timepoints in the window our data is in. Essentially, we are going to generate the best fit line from those parameters and the 95% confidence interval so that we can plot it with the original datapoints.

```{r}
dphz_grid <- tibble(time = seq(0.4, max(df_dphz_swv$time), length.out = 1000))

dphz_coefs <- left_join(dphz_nls %>% filter(term == 'b'),
                      dphz_nls %>% filter(term == 'a'), 
                      by = c('exp','run'), suffix = c('_b','_a'))

dphz_grid <- left_join(df_dphz_swv %>% filter(reactor == 'transfer') %>% group_by(reactor, exp, run, exp_id, run_id) %>% summarise(),
                       dphz_grid %>% mutate(reactor = 'transfer'), by = c('reactor'))

dphz_grid_coef <- left_join(dphz_grid, dphz_coefs, by  = c('exp','run'))

dphz_preds <- dphz_grid_coef %>% 
  mutate(pred = estimate_b * (time^-0.5) + estimate_a) %>% 
  mutate(pred_low = conf.low_b* (time^-0.5) + conf.low_a) %>% 
  mutate(pred_high = conf.high_b* (time^-0.5) + conf.high_a) 

write_csv(dphz_preds, "phz2019_dPHZ_Dphys_preds.csv")

# print simple results here
dphz_preds %>% 
  select(reactor,exp, run, time, pred, pred_high, pred_low, estimate_b, 
         conf.low_b, conf.high_b, estimate_a, conf.low_a, conf.high_a) %>% 
  head() %>% kable(digits = 10) %>% kable_styling() %>% scroll_box(height = '300px')
```
<br>

Now they we have those predictions we can plot everything together:

```{r}
plot_decays_dphz <- ggplot(dphz_preds, aes(x = time, y = pred)) + 
  geom_ribbon(aes(ymin = pred_low, ymax = pred_high), fill = 'light gray') +
  geom_path(linetype = 2)+
  geom_point(data =df_dphz_swv %>% filter(reactor == 'transfer'), aes(x = time, y = signal) , shape = 21)+
  facet_wrap(exp_id~run_id, scale = 'free')

plot_decays_dphz_styled <- plot_decays_dphz + 
  scale_y_continuous(labels = nA_label) +  
  labs(y = expression(I[swv]~(nA)), x = 'Time (min)', title = '∆phz*')

plot_decays_dphz_styled
  
```


# Fig. S6D

Now let's plot the SWV peak current decays for the blank IDA. This data comes from the blank IDA processing notebook.

```{r}

df_blank_swv <- read_csv("../../../processing/processed_data/phz_eDNA_2019_swv_blank_tran_time_signals.csv") %>% 
  filter(reactor == 'transfer' & PHZadded != '10uM') %>% 
  mutate(PHZadded = fct_relevel(PHZadded, c('25uM','50uM','75uM','100uM') ))

ggplot(df_blank_swv, aes(x = time, y = signal)) + 
  geom_point(shape = 21) + facet_wrap(~PHZadded, scales = 'free')
```

Now we will fit each of these decays. We will go ahead and save these coefficients as a csv, so that we can use them to calculate Dphys values in main figure 6.

```{r}
blank_nls <- df_blank_swv%>%
  group_by(PHZadded) %>% 
  do(
    tidy(
      nls(data = ., formula =  signal ~ b * (time)^-0.5 + a, start = c(b = 0.1, a = 1e-07) ),
      conf.int = T
    )
  ) %>% 
  arrange(desc(term))

# write results to csv for fig 6.
write_csv(blank_nls, "phz2019_blank_Dphys_nls_coefs.csv")

# print results here
blank_nls %>% kable(digits = 10) %>% kable_styling() %>% scroll_box(height = '300px')

```
<br>

Now we are going to take those coefficient estimates and predict datapoints that those values would generate at 1000 timepoints in the window our data is in. Essentially, we are going to generate the best fit line from those parameters and the 95% confidence interval so that we can plot it with the original datapoints.

```{r}
blank_grid <- tibble(time = seq(0.05, max(df_blank_swv$time), length.out = 1000))

blank_coefs <- left_join(blank_nls %>% filter(term == 'b'),
                      blank_nls %>% filter(term == 'a'), 
                      by = c('PHZadded'), suffix = c('_b','_a'))

blank_grid <- left_join(df_blank_swv %>% group_by(reactor, PHZadded) %>% summarise(),
                       blank_grid %>% mutate(reactor = 'transfer'), by = c('reactor'))

blank_grid_coef <- left_join(blank_grid, blank_coefs, by  = c('PHZadded'))

blank_preds <- blank_grid_coef %>% 
  mutate(pred = estimate_b * (time^-0.5) + estimate_a) %>% 
  mutate(pred_low = conf.low_b* (time^-0.5) + conf.low_a) %>% 
  mutate(pred_high = conf.high_b* (time^-0.5) + conf.high_a) 

write_csv(blank_preds, "phz2019_blank_Dphys_preds.csv")

# print simple results here
blank_preds %>% 
  select(reactor,PHZadded, time, pred, pred_high, pred_low, estimate_b, 
         conf.low_b, conf.high_b, estimate_a, conf.low_a, conf.high_a) %>% 
  head() %>% kable(digits = 7) %>% kable_styling() %>% scroll_box(height = '300px')
```
<br>

Now they we have those predictions we can plot everything together:

```{r}
plot_decays_blank <- ggplot(blank_preds, aes(x = time, y = pred)) + 
  geom_ribbon(aes(ymin = pred_low, ymax = pred_high), fill = 'light gray') +
  geom_path(linetype = 2)+
  geom_point(data =df_blank_swv, aes(x = time, y = signal) , shape = 21)+
  facet_wrap(~PHZadded, scale = 'free')

plot_decays_blank_styled <- plot_decays_blank+
    scale_y_continuous(labels = nA_label) +  
  labs(y = expression(I[swv]~(nA)), x = 'Time (min)', title = 'Blank')

plot_decays_blank_styled
```

# Create figure

```{r}
theme_set(theme_figure())

top_panel <- plot_grid(plot_dap_combo_styled + guides(fill = F), plot_decay_combo_styled + guides(fill = F, color= F))

fig_s6 <- plot_grid(top_panel,plot_swv_gc_dphz_styled, plot_decays_dphz_styled, plot_decays_blank_styled, 
                    ncol = 1, rel_heights = c(0.5, 1,1,1))

fig_s6

save_plot("../../../../figures/supplement/phz2019_Fig_S6.pdf", fig_s6, base_height = 10, base_width = 7)
```

-----

```{r}
sessionInfo()
```
