---
title: "Figure S3"
subtitle: 'Extracellular DNA promotes efficient extracellular electron transfer by pyocyanin in *Pseudomonas aeruginosa* biofilms.'
author: 'Scott H. Saunders, Edmund C.M. Tse, Matthew D. Yates, Fernanda Jiménez-Otero, Jacqueline K. Barton, Leonard M. Tender and Dianne K. Newman'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
    keep_md: true
---

--------

# Notes

Fig S1A is a set of images.

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(broom)
library(modelr)

# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)

# Load plotting tools
source("../../../tools/plotting_tools.R")


#Modify the plot theme
theme_set(theme_notebook())
```

# Fig. S1B

```{r}
etbr_data <- read_csv("../../../../data/Spectroscopy/phz_etbr_competitive_binding.csv") %>% 
  gather(key = 'rep', value = 'fluor', -phz_conc, -phz_mol)

# Plot layout
etbr_plot <- ggplot(etbr_data %>% filter(phz_conc>0), aes(x = phz_conc, y = fluor)) + 
  geom_point(shape = 21) + 
  geom_smooth(color = 'black', linetype = 2) + 
  facet_wrap(~phz_mol, scales = 'free', ncol = 1) + 
  ylim(0,2000)

# Plot styling
etbr_plot_styled <- etbr_plot +
  labs(x = expression("Added phenazine"~(mu*M)), y = 'Ethidium Fluorescence') + 
  scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c("0.1", "1", "10", "100", "1000"))
  
etbr_plot_styled


etbr_data %>% group_by(phz_mol, phz_conc) %>% summarise(mean_fluor = mean(fluor)) 
```

```{r}
etbr_model_pyo <- nls(data = etbr_data %>% 
                        filter(phz_conc>0 & phz_mol=='PYO'), 
                      formula = fluor~max - (max*(phz_conc^hill)/((ic50^hill)+(phz_conc^hill))),
                      start=c(max=1500,hill=1,ic50=300))

etbr_model_pyo_CI <- tidy(etbr_model_pyo, conf.int = T) %>% filter(term=='ic50')




etbr_model_pcn <- nls(data = etbr_data %>% 
                        filter(phz_conc>0 & phz_mol=='PCN'), 
                      formula = fluor~max - (max*(phz_conc^hill)/((ic50^hill)+(phz_conc^hill))),
                      start=c(max=1500,hill=1,ic50=300))

#confint(etbr_model_pcn)

summary(etbr_model_pyo)

etbr_model_pcn_CI <- tidy(etbr_model_pcn) %>% filter(term=='ic50') %>% 
  mutate(conf.low = estimate - (1.96 * std.error)) %>% 
  mutate(conf.high = estimate + (1.96 * std.error))

etbr_pred_pyo <- etbr_data %>% filter(phz_mol=='PYO') %>% add_predictions(etbr_model_pyo)

etbr_pred_pyo <- etbr_data %>% filter(phz_mol=='PYO') %>% 
  mutate(max = coef(etbr_model_pyo)[1], hill = coef(etbr_model_pyo)[2], ic50 = coef(etbr_model_pyo)[3]) %>% 
  mutate(ic50_high = etbr_model_pyo_CI$conf.high,ic50_low = etbr_model_pyo_CI$conf.low) %>% 
  mutate(pred = max - (max*(phz_conc^hill)/((ic50^hill)+(phz_conc^hill)))) %>% 
  mutate(pred_high = max - (max*(phz_conc^hill)/((ic50_high^hill)+(phz_conc^hill)))) %>% 
  mutate(pred_low = max - (max*(phz_conc^hill)/((ic50_low^hill)+(phz_conc^hill))))

etbr_pred_pcn <- etbr_data %>% filter(phz_mol=='PCN') %>% 
  mutate(max = coef(etbr_model_pcn)[1], hill = coef(etbr_model_pcn)[2], ic50 = coef(etbr_model_pcn)[3]) %>% 
  mutate(ic50_high = etbr_model_pcn_CI$conf.high,ic50_low = etbr_model_pcn_CI$conf.low) %>% 
  mutate(pred = max - (max*(phz_conc^hill)/((ic50^hill)+(phz_conc^hill)))) %>% 
  mutate(pred_high = max - (max*(phz_conc^hill)/((ic50_high^hill)+(phz_conc^hill)))) %>% 
  mutate(pred_low = max - (max*(phz_conc^hill)/((ic50_low^hill)+(phz_conc^hill))))

#etbr_pred_pcn <- etbr_data %>% filter(phz_mol=='PCN') %>% add_predictions(etbr_model_pcn)

ggplot(etbr_pred_pyo, aes(x = phz_conc, y = fluor)) + 
  geom_ribbon(aes(ymin = pred_low, ymax = pred_high), fill = 'light gray')+
  geom_line(aes(y = pred), linetype = 2) + 
  geom_point(shape = 21) + 
  scale_x_log10()

ggplot(etbr_pred_pcn, aes(x = phz_conc, y = fluor)) + 
  geom_ribbon(aes(ymin = pred_low, ymax = pred_high), fill = 'light gray')+
  geom_line(aes(y = pred), linetype = 2) + 
  geom_point(shape = 21) + 
  scale_x_log10()

ggplot(etbr_data %>% filter(phz_conc>0 & phz_mol=='PYO') %>% mutate(norm_fluor = -fluor + 1500), 
       aes(x = phz_conc, y = norm_fluor)) + 
  geom_point(shape = 21) + 
  scale_x_log10()

```

```{r}
etbr_preds = bind_rows(etbr_pred_pyo, etbr_pred_pcn, etbr_data %>% filter(phz_mol=='PCA'))

etbr_plot <- ggplot(etbr_preds %>% filter(phz_conc>=1), aes(x = phz_conc, y = fluor)) + 
  geom_ribbon(aes(ymin = pred_low, ymax = pred_high), fill = 'light gray')+
  geom_line(aes(y = pred), linetype = 2, size = 0.5) + 
  geom_point(shape = 21, size = 1) + 
  facet_wrap(~phz_mol, scales = 'free', ncol = 1)

etbr_plot_styled <- etbr_plot +
  labs(x = expression("Added phenazine"~(mu*M)), y = 'Ethidium Fluorescence') + 
  scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c("0.1", "1", "10", "100", "1000"))
  
etbr_plot_styled
```

```{r}

etbr_kd = 1 #uM

etbr_conc = 5 #uM 


```

$$K_i = \frac{IC_{50}}{1+\frac{[EtBr]}{K_d}}$$

$$K_i = \frac{IC_{50}}{6}$$

For PCN Ki = 612 / 6 = 102uM
PCN low Ki = 450 / 6 = 75
PCN high Ki = 774 / 6 = 129

For PYO Ki = 275 / 6 = 45.8 uM
PYO low Ki = 214 / 6 = 35.8
PYO high Ki = 364 / 6 = 60.7

```{r}
#save_plot("../figures/03_14_19_fig_supp_etbr.pdf", etbr_plot_styled ,base_width = 4, base_height = 6 )

```

```{r}
mst_data <- read_csv("../../../../data/Spectroscopy/MST_data_phz_DNA.csv") %>% 
  mutate(norm_resp = Response - 640)

ggplot(mst_data, aes(x = Dose, y = norm_resp)) + 
  geom_point() + 
  geom_smooth(method='nls',
              formula=y~max*(x^hill)/((Kd^hill)+(x^hill)),
              method.args=list(start=c(max=400,hill=1,Kd=100)),
              se=F, color = 'black', linetype = 'dashed', size = 0.5) +
  facet_wrap(~phz_mol, scales = 'free', ncol=1)

model_pyo <- nls(data = mst_data %>% filter(phz_mol=='PYO') ,formula = norm_resp~max*(Dose^hill)/((Kd^hill)+(Dose^hill)),start=c(max=400,hill=1,Kd=100) )
  
model_pcn <- nls(data = mst_data %>% filter(phz_mol=='PCN') ,formula = norm_resp~max*(Dose^hill)/((Kd^hill)+(Dose^hill)),start=c(max=400,hill=1,Kd=100) )

confint_pyo <- tidy(model_pyo, conf.int = T) %>% filter(term=='Kd')

summary(model_pcn)

confint_pcn <- tidy(model_pcn) %>% filter(term=='Kd') %>% 
  mutate(conf.low = estimate - (1.96*std.error), conf.high = estimate + (1.96*std.error))
  
pred_pyo <- mst_data %>% filter(phz_mol=='PYO') %>% 
  mutate(max = coef(model_pyo)[1], hill = coef(model_pyo)[2], kd = coef(model_pyo)[3]) %>% 
  mutate(kd_low = confint_pyo$conf.low,kd_high = confint_pyo$conf.high ) %>% 
  mutate(pred = max*(Dose^hill)/((kd^hill)+(Dose^hill))) %>% 
  mutate(pred_low = max*(Dose^hill)/((kd_low^hill)+(Dose^hill))) %>% 
  mutate(pred_high = max*(Dose^hill)/((kd_high^hill)+(Dose^hill))) 

pred_pcn <- mst_data %>% filter(phz_mol=='PCN') %>% 
  mutate(max = coef(model_pcn)[1], hill = coef(model_pcn)[2], kd = coef(model_pcn)[3]) %>% 
  mutate(kd_low = confint_pcn$conf.low,kd_high = confint_pcn$conf.high ) %>% 
  mutate(pred = max*(Dose^hill)/((kd^hill)+(Dose^hill))) %>% 
  mutate(pred_low = max*(Dose^hill)/((kd_low^hill)+(Dose^hill))) %>% 
  mutate(pred_high = max*(Dose^hill)/((kd_high^hill)+(Dose^hill))) 

preds <- bind_rows(pred_pyo,pred_pcn,mst_data %>% filter(phz_mol=='PCA'))

# Plot layout
mst_plot <- ggplot(preds, aes(x = Dose, y = norm_resp)) + 
  geom_ribbon(aes(ymin = pred_low, ymax = pred_high), fill = 'light gray')+
  geom_line(aes(y = pred), linetype = 2, size = 0.5) +
  geom_point(shape = 21, size = 1) + 
  facet_wrap(~phz_mol, scales = 'free',ncol=1)+ 
  scale_x_log10()

# Plot styling
mst_plot_styled <- mst_plot +
  labs(x = expression("Added phenazine"~(mu*M)), y = "Thermophoresis response")

mst_plot_styled
```

Kd for PCN = 133.6 uM
Kd for PYO = 536.4 uM (but doesn't saturate)

# PHZred fluorescence


## Import data

```{r}
df_360 <- read_csv("../../../../data/Spectroscopy/2019_10_09_PHZred_top_360-460.csv") %>% 
  gather(key = 'column', value = 'intensity', -row)

df_485 <- read_csv("../../../../data/Spectroscopy/2019_10_09_PHZred_top_485-528.csv") %>% 
  gather(key = 'column', value = 'intensity', -row)

df_meta <- read_csv("../../../../data/Spectroscopy/2019_10_09_PHZred_well_metadata.csv") %>% 
  gather(key = 'column', value = 'id', -row)

df_360_meta <- left_join(df_360, df_meta, by = c('row','column')) %>% 
  mutate(dna = case_when(
    column %in% c(1,7) ~ 0,
    column %in% c(2,8) ~ 17.5,
    column %in% c(3,9) ~ 35,
    column %in% c(4,10) ~ 70,
    column %in% c(5,11) ~ 140,
    column %in% c(6,12) ~ 280,
  ))

df_485_meta <- left_join(df_485, df_meta, by = c('row','column')) %>% 
    mutate(dna = case_when(
    column %in% c(1,7) ~ 0,
    column %in% c(2,8) ~ 17.5,
    column %in% c(3,9) ~ 35,
    column %in% c(4,10) ~ 70,
    column %in% c(5,11) ~ 140,
    column %in% c(6,12) ~ 280,
  ))
```


## 360nm ex

```{r}
ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity)) + geom_point() + facet_wrap(~id, scales = 'free')

ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity, color = id)) + geom_point() + geom_smooth(method = 'lm')
```

```{r}
#ggplot(df_360_meta %>% filter(id=='empty'), aes(x = as.integer(column), y = intensity, color = row)) + geom_point()

#ggplot(df_360_meta %>% filter(id=='PBS'), aes(x = as.integer(column), y = intensity, color = row)) + geom_point()

#ggplot(df_360_meta %>% filter(id %in% c('empty','PBS')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()
```

```{r}
ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PCA')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()

ggplot(df_360_meta %>% filter(id %in% c('empty','PBS','PCN')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()

ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PYO')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()

```

```{r}
#ggplot(df_360_meta %>% filter(id %in% c('PYO')), aes(x = dna, y = intensity, color = id)) + geom_point()

#ggplot(df_360_meta %>% filter(id %in% c('PBS')), aes(x = dna, y = intensity, color = id)) + geom_jitter(shape = 21)


```

## 485nm ex

```{r}
ggplot(df_485_meta %>% filter(id %in% c('control','empty','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity)) + geom_point() + facet_wrap(~id, scales = 'free')

ggplot(df_485_meta %>% filter(id %in% c('control','empty','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity, color = id)) + geom_point() + geom_smooth(method = 'lm')
```

## Figure 

```{r}

df_fig <- bind_rows(df_360_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PYO','PBS','control')) %>% mutate(ex = '360'),
          df_485_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PCA','PBS','control')) %>% mutate(ex = '485')
          )

df_fig <- df_fig %>% group_by(id,ex) %>% mutate(max = max(mean)) %>% mutate(norm_int = intensity / max, norm_mean = mean / max)

#ggplot(df_fig, aes(x = dna, y = norm_int, color = id, shape = ex)) + geom_point()

#ggplot(df_fig, aes(x = dna, y = norm_mean, color = id, shape = ex)) + geom_point()

#ggplot(df_fig, aes(x = dna, y = norm_mean, color = id, shape = ex, linetype = ex)) + geom_point(aes(y = norm_int)) + geom_line() + ylim(0,NA)

ggplot(df_fig, aes(x = dna, y = norm_mean, color = id)) + geom_point(aes(y = norm_int)) + geom_line() + ylim(0,NA) + facet_wrap(~ex)

#ggplot(df_fig %>% filter(id != 'PBS'), aes(x = dna, y = norm_mean, color = id)) + geom_jitter(aes(y = norm_int)) + geom_line() + ylim(0,NA)

facet_names <- list(
  '360'="PYO",
  '485'="PCA"
)

facet_labeller <- function(variable,value){
  return(facet_names[value])
}

ggplot(df_fig %>% filter(id != 'PBS'), aes(x = dna, y = norm_mean, color = id)) + geom_jitter(aes(y = norm_int), shape = 21) + geom_line() + 
  facet_wrap(~ex, labeller = facet_labeller, scales = 'free') + 
  scale_color_manual(values = c('light gray','black','black')) + scale_y_continuous(labels = scales::percent, limits = c(0,NA)) + 
  labs(x = 'bp DNA (uM)', y = 'Normalized fluorescence')
```

Trying to make figure to match etbr and mst data:

```{r}
df_fig_2 <- bind_rows(df_360_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PYO','control')) %>% mutate(ex = '360_PYO'),
          df_485_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PCA','control')) %>% mutate(ex = '485_PCA'),
          df_485_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PCN','PBS')) %>% mutate(ex = '485_PCN')
          )

df_fig_2 <- df_fig_2 %>% group_by(id,ex) %>% mutate(max = max(mean)) %>% mutate(norm_int = intensity / max, norm_mean = mean / max) %>% 
  ungroup() %>% 
  mutate(ex = fct_relevel(ex,'485_PCA','485_PCN','360_PYO' ))

facet_names <- list(
    '485_PCA'="PCA",'485_PCN'="PCN",'360_PYO'="PYO"
)

facet_labeller <- function(variable,value){
  return(facet_names[value])
}

#df_fig_2$ex = factor(df_fig_2$ex, levels=c('485_PCA','485_PCN','360_PYO'))

#scale_color_manual(values = c('light gray','black','black')) + 

plot_phz_red <- ggplot(df_fig_2, aes(x = dna, y = norm_mean, color = id)) + geom_point(aes(y = norm_int), shape = 21, size = 1) + geom_line(size = 0.5) + 
  facet_wrap(~ex, scales = 'free', ncol = 3) + 
  scale_y_continuous(labels = scales::percent, limits = c(0,1.25)) + 
  labs(x = 'bp DNA (uM)', y = 'Normalized fluorescence') + scale_color_manual(values = c('light gray','light gray','black','black','black')) + guides(color = F) + xlim(0,300)


plot_phz_red
```
Conclusions: does not seem to be signficant fluorescence change compared to empty wells. Note that there is nonzero variation between wells in plate reader. Reduced phenazines do not bind strongly to DNA.

Caveats: 90uM phenazine may be well below the Kd, so this only provides some evidence that there's no binding at this concentration of phenazine.

# Create Figure

```{r}
binding_plots <- plot_grid(etbr_plot_styled,mst_plot_styled, ncol=2, scale = 0.95, align = 'hv', axis = 'tblr', labels = c("A","B"))

bottom_plot <- plot_grid(plot_phz_red, ncol=1, align = 'hv', axis = 'tblr', labels = c('C'), scale = 0.95)

fig_s3 <- plot_grid(binding_plots,bottom_plot, ncol=1, align = 'hv', axis = 'tblr', rel_heights = c(2,1))

fig_s3

save_plot("../../../../figures/supplement/phz2019_Fig_S3.pdf", fig_s3, base_width = 7, base_height = 6)

```

----

```{r}
sessionInfo()
```